Complex mecanisms linking biomass stability and total biomass with species richness and network structure in stream ecosystems

### 2020-04-27 manuscript/bef_stability/manuscript.Rmd
Update manuscript + appendix

---
title:"""
  Complex mecanisms linking biomass stability and total biomass with
species richness and network structure in stream ecosystems 
"""
author: Alain Danet, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(7, 7),
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE
)

@@

Species richness has been reported to increase productivity
[@tilman_productivity_1996, @grace_integrative_2016] through complementarity
between species in resource exploitation and selection effects
[@loreau_partioning_2001, @cardinale_species_2002, @fargione_from_2007].
Empirical assessment of the effect of species richness on total biomass showed negative
[@grace_integrative_2016] and positive relationships [@hector_plant_1999,
@tilman_biodiversity_2014]. @grace_integrative_2016 showed that the results
depend on the refinement of the mecanisms involved in the analysis.

@@

---
#- Research questions:
  #- What are the effects of the community structure on total biomass and stability ? 
    #1. What are the effects of species richness on total biomass and stability ?
      #- effect of species richness
      #- effect of trophic group
    #2. What are the effects of the community structure on total biomass and stability ?
      #- What is the relative effect of the environment relative to the community
	#structure ?
      #- What is the effect of the community structure on the different
	#parts of the stability + total biomass?
	#- Effect of horizontal and vertical diversity
---

@@

---
#- Environment can drive stability and biomass:
  #- forcing: environmental forcing on communities 
  #- can lead to misleading effect: 
---

@@

@@

---
#- OFB
  #- French long term protocol
  #- standardized
#- Electric fishing:
  #- stuned fish
  #- identification of the fishes
  #- fish counted  
  #- length of the fishes in mm by lot (4):
    #- I: big fishes measured individually
    #- S/L: fishes are grouped by species and by size 
      #- 30 random fishes were measured in each lot
      #- simulate the size of all the fishes of the lot:
	#- normal distribution truncated at 95% (?)
	#- mean and sd calculted from the sample of the lot 
    #- G: little fishes are grouped by species
      #- min and max size by lot 
      #- simulate the size of all the fishes of the lot:
	#- normal distribution truncated at 95% (?)
	#- mean computed as ???  
	#- se computed as ??? 
    #- N:
#- Protocols type:
  #- complete
    #- shallow river
  #- partial:
    #- deep and large river
    #- partial over bank
    #- partial by point
    #- fusion of the two Protocols when "by point" all the point were done
      #over bank
---

@@

Species richness was computed as the total species richness over time in a given
station. To normalize by sampling effort, we reported total species richness to
the sum of surface sampled over the time period, it was expressed in number of
species by square meter.

@@

---
#- Diet data: 
  #- 7? species over 57 had a piscivorous stage
  #- 3? species had a strict piscivorous stage
  #- ontogenic diet shift based on size
#- 7 Resource nodes:
  #- based on diet database
  #- no information about their biomass
#- Size class:
  #- quantile: 9 size class per species
---

---
#- Interaction:
  #- Fish-resources and resources: based on diet data 
  #- Fish-fish:
    #1. diet data: is piscivorous?
    #2. allometric relationship: predation range from .03 to .45 of body size
---

Network inference was performed in two steps: determine trophic species,  build
a metaweb, subtract the metaweb according to the trophic species present in
community. Diet information were obtained from database and literature. Over
the 57 fish species, only seven had a piscivorous life stage and only three had
as strict piscivorous life stage. Fish species were largely omnivorous. Because
species have ontogenic diet shift, we created nine size classes by species, one
size class of one species constituting one trophic species. The bounds between
size classes were equidistant (percentile method). The metaweb described the
potential trophic interactions between the $`r 57*9`$ ($57\times9$) trophic species. 
We had seven resource nodes which were present in the diet of the trophic
species namely detritus, biofilm, phytoplankton, zooplankton, macrophages,
phytobenthos and zoobenthos. A trophic species belonged to a life stage if the
median of its size range fell in the range of the life stage. In turn, this
trophic species has feeding interactions with the resource nodes of its life stage.
Interactions between resource nodes are determined by literature. For example,
phytobenthos fed on detritus, while zooplankton fed on zooplankton and
phytoplankton. Fish-fish trophic interactions were determined by allometric
relationships as in many studies
[@brose_predator_2019, @poisot_structure_2016, @schneider_body_2012].
For a given trophic species, we first determined if its
size class end up in a life stage containing piscivorous diet. If it did, we
determined the prey size range of the trophic species from 3% to 45% of the
median of its size range. A feeding interaction was set for each trophic species
whose median of its size class fell into the prey range of the predator trophic
species. To sum up, the metaweb was obtained with allometric feeding
interactions (1) between trophic species, (2) between trophic species and
resource nodes and (3) between resource nodes.

The local networks (i.e. one network for each fish community sampling) was
obtained by determining the trophic species of each fish individual in the
community and interactions were set according to the metaweb. All the resource
nodes were assumed to be present in each local network. Finally, the
structure of the local networks varied according the presence of the trophic
species. The inferred interactions represent potential interactions (i.e.
not measured interactions) and are binary (i.e. presence or absence of feeding
interactions).

We computed the connectance ($C = L/N^2$, L and N being resp. the number of
links and of nodes) and the average trophic level of each network, the average
trophic level was weighted the biomass of each node (excluding resource nodes
for which we had no biomass) ($\overline{T} = \sum_i^N{B_iT_i \times
\sum_i^N{B_i}^{-1}}$, $T_i$ and $B_i$ being resp. the trophic level and the
biomass of the node $i$). In order to have connectance at the species level
(i.e. not a trophic species level), we merge the trophic species of the network
by species and thereby aggregating the link by species.

## Community

---
#- Biomass:
  #- infered by allometric law 
  #- approximated by $B = 0.01 \times (l^3.03)$
---

@@

# Compute stability, network metrics, etc with ;he new datasets
#debug(summarise_network_over_time)
##myload(network_metrics, dir = mypath("data", "classes"))
#network_metrics %>%
  #left_join(op_analysis, by = "opcod") %>%
  #select(opcod, station, composition) %>%
  #unnest(composition) %>%
  #group_by(station, troph_group) %>%
  #summarise(richness_tot = length(unique(species)))
com_data <- compute_community_temporal_analysis(.op = op_analysis_bbb,
 type_network_metrics = "classes" 
)
# Compute sem dataset 
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]],
  hab_press = habitat_pressure,
  sync = com_data[["sync_std"]],
  nmds = NULL,
  basin = st_basin
  )

stab_div_troph <- com_data$tps_bm_troph %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group),
    log_rich = log10(richness_med),
    log_rich_std = log10(rich_std_med),
    log_rich_tot = log10(richness_tot),
    log_rich_tot_std = log10(rich_tot_std),
    log_bm = log10(biomass_med),
    log_bm_std = log10(bm_std_med),
    log_stab = log10(biomass_stab),
    log_stab_std = log10(bm_std_stab)
    ) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med, richness_tot, log_rich, log_rich_std, log_rich_tot, log_rich_tot_std, log_stab, log_stab_std, log_bm, log_bm_std, biomass_med, bm_std_med, richness_tot, rich_tot_std) %>%
  dplyr::left_join(st_basin, by = "station")
```

@@

---
#- Collected during fishing operation:
  #- River size (width and depth)
  #- Altitude and slope
  #- distance to source
#- Independant dataset from naiades: 
  #- flow, temperature, DBO
    #1. moving window average
    #2. annual average 
    #3. interpolation along rivers (OpenStars + SSN package)
---

Habitat description were obtained from fish sampling data. We kept variables
describing the ecosystem size namely the stream width and depth, the ecosystem
position (altitude and slope). Physico-chemical variables were obtained from
naiades (http://www.naiades.eaufrance.fr/)[http://www.naiades.eaufrance.fr/].
We chose flow and temperature as physical variables to describe streams as flow
impose swimming constraints which filter fish species that can establish [REF].
Water temperature is linked to physiological constraints namely thermal
tolerance and oxygen concentration tolerance [REF] since the level of water
saturation in oxygen is inversely linked to water temperature. We chose
Biochemical Oxygen Demand (BOD) for a proxy of enrichment in the ecosystem [REF
mccann]. BOD is also inversely linked to oxygen concentration in water.
The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD). For each recording
station, we applied a moving window averaging (more details in supplementary
material) to get rid of seasonal effects and took the annual mean (i.e. one
value by year and by station. As the environmental recording stations were not
located at the same place than the fish stations, we interpolated data with
spatial statistical network model [@isaak_applications_2014] implemented in the 
`SSN` R package [@hoef_ssn:_2014]. The interpolation procedure is detailed
in supplementary material. For each fish station, environmental variables were
matched with the year of fish sampling (except for water temperature for which the 
record began in 2006). Then, the average temperature, flow, BDO, river width
and depth were computed as well as coefficient of variation ($CV = \sigma
\mu^{-1}$). As habitat and environmental variables are often colinear,
a principal component analysis (PCA) was performed on the scaled variables. The
PCA explained % of the variance. An axis rotation was performed to maximize the
representation of the variables by the principal components
(Fig. \@ref{fig:pca-rotated}). We kept five axis, representing respectively the
(1) river size (river width & depth), (2) the altitude, slope and decreasing
average temperature, (3) average enrichment (BDO) and CV of flow, (4) river
size CV (CV of river width and depth) and (5) CV of enrichment (CV of BDO).
Additionally, we logged the axis one, two and three to normalize their
distribution and we reversed the third axis to get increasing average
temperature with increasing values of the axis.

@@

pca_rotated <- compute_rotated_pca(
  .data = pca_pressure[, !names(pca_pressure) %in% "station"],
  naxis = 5 
)
```

@@

### Relationship between stability and species richness

$$\begin{aligned}
\log_{10}S_i &= \beta_0 + \beta_R\log_{10}x_{Ri} + \epsilon_{ij}\\
S_i &= 10^{\beta_0 + \beta_R\log_{10}x_{Ri} + \epsilon_{ij}}\\
S_i &= 10^{\beta_0} \times 10^{\beta_R\log_{10}x_{Ri}} \times10^{\epsilon_{ij}}\\
S_i &= 10^{\beta_{0}} \times x_{Ri}10^{\beta_{R}} \times 10^{\epsilon_{ij}}
\end{aligned}$$

with $i$ being the station $i$.

@@

```{r mod-div-total}
stab_div_mod <- lmer(
  data = sem_data,
  control = ctrl_lmer,
  formula = log_stab_std ~ log_rich_tot_std + (1 | basin)
)
stopifnot(!isSingular(stab_div_mod, tol = 1e-05))
stab_div_boot <- bootMer(stab_div_mod, mySumm2, nsim = 100)
stab_div_ci <- confint(stab_div_boot)
#stab_div_colin <- car::vif(stab_div_mod) not use if one response
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

@@

```{r mod-div-troph, fig.show = "hide"}
stab_troph_mod <- lmer(data = na.omit(stab_div_troph),
  formula = log_stab_std ~ log_rich_tot_std + troph_group + log_rich_tot_std:troph_group + (1 | basin)
  #control = ctrl
)

@@

## get coeff
#coef_troph <- map(list(stab = stab_troph_mod, bm = bm_troph_mod), fixed.effects)
```
```{r stab-div-troph}
stab_div_by_troph <- compute_stat_troph_rich(
  .data = stab_div_troph,
  resp = "log_stab_std"
)

bm_div_by_troph <- compute_stat_troph_rich(
  .data = stab_div_troph,
  resp = "log_bm_std"
)
troph_perm_test <- map(c(stab = "log_stab_std", bm = "log_bm_std"),
  ~test_troph(.data = stab_div_troph, var = .x)
  )
troph_perm_test$bm$obs
troph_perm_test$stab$obs
troph_perm_test$bm$p_val
troph_perm_test$stab$p_val

@@

We computed Structural Equation Modeling (SEM) to unravel mechanisms underlying link
between environment, community structure and both biomass stability and total
biomass. Why SEM?

SEM enables to reason in terms of causal relationship between components. We
aimed to understand the mechanisms behind the negative relationship biomass
stability and species richness and the positive relationship between total
biomass and species richness.

We built a metamodel to set up the nodes and the links between the different
compartment of the SEM. Basically, we tested the links between environmental
conditions, community structure and stability. Theory in ecology predicts that
community structure influences stability. In turn, environment influences
community structure but also directly stability.

@@

To characterize environment, we retained five axis of PCA on habitat and
climatic variables which represent different types of forcing (constant and
variability). The two first axis (RC1 and RC2) were linked to the physical
structure and position of the river and the altitutinal gradient (+temperature).
The third axis (RC3) was linked to enrichment (Organic Biological Demand) and CV
of annual flow. The fourth and fifth (RC4 and RC5) axis were associated to
environmental variability, respectively to CV of width and depth of the river
and CV of enrichment.

The community structure was characterized by total species richness,
median average weighted trophic level and median connectance.
Total species richness was the number of unique fish species that were observed
at a given station over the whole sampling. Both median
connectance and average trophic level were just computed as the median over the
temporal period. For the details of the computation of the trophic networks,
see Network inference part.

@@

---
# Environment to community attributes and stability  
---
We linked environmental variables both to community structure and stability
components. Rooted in assembly theory, environment filters species that can
establish in a community and thus affect community structure. (+ add specific
hypothesis about physical and climate). (1) We hypothesized that enrichment
(characterized by DBO) and mean annual temperature will increase species turnover
[@friedpetersen_drivers_2020]. (2) We hypothesized environmental forcing can also
decrease biomass stability directly without going by community structure. First,
instability of habitat and environmental factors may in turn drive biomass
instability [@hansen_climate_2013]. Secondly, average environmental conditions
such as enrichment can increase directly CV by increasing species biomass
fluctuations [@tabi_warming_2019]. Increasing temperature was also hypothesized
to decrease stability through starvation resulting from high metabolic activity
[@tabi_warming_2019].

---
# Species richness to community attributes 
---
To reveal the mechanisms behind the relationship between species richness and
biomass stability, causal links were set from species richness to other
community attributes. (5) We hypothesized also that species richness
decreased connectance following results from @thebault_stability_2010;
@winemiller_must_1989. But theoretically, some ecological networks can show the
inverse relationship [@winemiller_must_1989]. (6) We hypothesized also that species richness
will modulate average weighted trophic level because of food web assembly
processes. If *species richness increased and that new species are placed at
random in the network, we expected a positive relationship between average
trophic level and species richness because individuals of the higher trophic
level have a higher body mass [to think again]*. Finally, we set up causal
relationships from community attributes to stability components.

---
#  Community attributes to stability
---
We hypothesized that (7) species richness decreased synchrony between species
[REF] and increased $\overline{CV_{sp}}$ [@tilman_biodiversity_2006] *to
complete*. (8a) We expected that average trophic level had a positive or
negative relationship with synchrony, bottom-up effects should result in
positive whereas top-down in negative relationship. We expected a (8b) negative
relationship between average trophic level and $\overline{CV_{sp}}$ if predators
stabilised population
fluctuations or if the increasing part of biomass in the predators decreased
$\overline{CV_{sp}}$ because of their higher body mass (if mean increase faster than
variance). (9a) Connectance was expected to decrease or increase $\overline{CV_{sp}}$
and (9b) synchrony because of top down effects or bottom up effects.

@@

library(DiagrammeR)
p_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "stab_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*30/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = c(from = 0, to = 7.5),
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2")
)

p_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "bm_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*45/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = c(from = 0, to = 7.5),
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2")
)
```

@@

---
#- Fishing operations and station
  #- kept one protocol type by station
  #- kept one sampling operation per year (of the season the most sampled in the
    #time series)
  #- remove operations which sampling effort (length of river sampled) superior to average
    #plus or minus 30% of sd (by station)
  #- kept station that had 10 or more samplings, i.e. followed 10 years or more
---

Only one protocol
type was kept by station. We kept only one sampling event by year to homogenize
the dataset [numbers: month of sampling, number of days bw samplings] (detailed
in Supplementary material).

@@

```{r sem-plot, fig.dim = c(10,10), fig.cap = sem_caption}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple

>>>>

---
title: Complex mecanisms linking biomass stability and total biomass with species richness and network structure in stream ecosystems 
author: Alain Danet, Maud Mouchet, Elisa Thebault and Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE
)

@@

Species richness has been reported to increase productivity
[@tilman_productivity_1996, @grace_integrative_2016] through complementarity
between species in resource exploitation and selection effects
[@loreau_partitioning_2001, @cardinale_species_2002, @fargione_selection_2007].
Empirical assessment of the effect of species richness on total biomass showed negative
[@grace_integrative_2016] and positive relationships
[@hector_plant_1999, @tilman_biodiversity_2014]. @grace_integrative_2016 showed
that the results depend on the refinement of the mecanisms involved in the
analysis.

@@

@@

@@

```{r}
matmet_cap <- paste0(
  "Mat&met figure"
)
```

```{r matmet, fig.cap = matmet_cap}
myload(p_st,temporal_network, dir = mypath("manuscript/bef_stability/figs"))
p_st <- p_st +
  theme(legend.position = "none")
# Network inference
net_method_path <- mypath("manuscript/bef_stability/figs",
  "network_inference.pdf")
net_method <- net_method_path %>%
  magick::image_read()
p_net_method <- cowplot::ggdraw() +
  cowplot::draw_image(net_method)

top <- plot_grid(p_st, p_net_method, ncol = 2, rel_widths = c(.8, 1), labels =
  "AUTO")
p_method <- plot_grid(top, temporal_network, nrow = 2, rel_heights = c(1, .3),
  labels = c("", "C"))
p_method
```

@@

@@

Species richness was computed as the total species richness over time in a given
station. To normalize by sampling effort, we reported species richness to
the sum of surface sampled over the time period, it was expressed in number of
species by square meter.

@@

To obtain the network structure of the communities, we needed to infer
ecological interactions between the components of the communities and thus build
the ecological network. As the interactions were not measured, we used trait
matching method to infer the presence or absence of an interaction. This method is
widely used to infer trophic interactions
[@brose_predator_2019, @poisot_structure_2016, @schneider_body_2012]. Indeed,
the existence of an trophic interaction between two species have been reported
to be largely dependant on their body size ratio [cascade model, @cohen_stochastic_1985, @warren_invertebrate_1987, @brose_predator_2019].

Additionally to body size, we used food diet data to infer trophic
interactions. The diet of stream fishes, as many organisms, shifts over lifespan.
For example, alvins feeds on plankton while adults may feed on algae or fishes.
As we got body size of each fish individuals (cf Biomass inference methods and
supplementary material), we are able to define trophic interactions beyond
species level. We thus divided each fish species into nine body size classes,
each class having the same body size range (percentile method). Each body size
class constituting a trophic species. We then infered trophic interactions
between the trophic species. At this stage, it is important to make clear that
we infered potential trophic interactions and not actual measured interactions.
We determined network structure in two steps: (1) build a metaweb decribing the
trophic interactions between all the possible trophic species, then (2)
determine network of a given community by subtracting the metaweb according to
the trophic species present in a given community.

### Metaweb

#### Ontogenic diet

Diet information were obtained from database and literature [REF edeline?]. The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had as strict piscivorous
life stage. Fish species were largely omnivorous.

#### Define trophic interactions

The metaweb described the potential trophic interactions between the $`r 57*9`$
($57\times9$) trophic species. We had seven resource nodes which were present
in the diet of the trophic species namely detritus, biofilm, phytoplankton,
zooplankton, macrophages, phytobenthos and zoobenthos.

A trophic species belonged to a life stage if the median of its body size range
fell in the range of the life stage. In turn, this trophic species has feeding
interactions with the resource nodes of its life stage. Interactions between
resource nodes were determined by literature. For example, phytobenthos fed on
detritus, while zooplankton fed on zooplankton and phytoplankton. Fish-fish
trophic interactions between a predator and a prey were determined by body size
ratio as in many studies
[@brose_predator_2019, @poisot_structure_2016, @schneider_body_2012].
For a given trophic species, we first determined if its size class end up in a
life stage containing piscivorous diet, i.e. if there was an overlap between
them. We were more liberal for piscivory determination to not miss potential
fish-fish trophic interactions which were the main interactions in our
community. If the trophic species was piscivorous, we determined its prey size range
from 3% to 45% of the median of its size range. A trophic interaction was set
for each trophic species whose median of its size class fell into the prey range
of the predator trophic species. To sum up, the metaweb was obtained with
allometric feeding interactions (1) between trophic species, (2) between trophic
species and resource nodes and (3) between resource nodes.

#### Obtain network of a community

The trophic network of a community (i.e. one network for each fish community
sampling) was obtained by determining the trophic species of each fish
individual in the community and interactions were set according to the metaweb.
All the resource nodes were assumed to be present in each local network.
Finally, the structure of the local networks varied according the presence of
the trophic species. The inferred interactions represent potential interactions
(i.e.  not measured interactions) and are binary (i.e. presence or absence of
feeding interactions).

#### Network structure

In each network, we computed two network metrics that have been commonly used to
characterize network structure. Connectance which describes the number of
interactions compared to the number of possible links,
 computed as ($C = L / N^2$, L and N being resp. the number of
links and of nodes). We then computed the average trophic level, the
average trophic level was weighted by the biomass of each trophic species
(excluding resource nodes for which we had no biomass) ($\overline{T} =
\sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, $T_i$ and $B_i$ being resp. the
trophic level and the biomass of the node $i$).

To sum up, we infered binary trophic networks at the scale of trophic species
(i.e. body size classes by species) to take in account ontogenic diet shifts and
that fish-fish trophic interaction depends on body size ratio. On the other
side, the community metrics were defined at the species level, (i.e. species
richness, average species CV and synchrony).

@@

# Compute stability, network metrics, etc with ;he new datasets
##myload(network_metrics, dir = mypath("data", "classes"))
#network_metrics %>%
  #left_join(op_analysis, by = "opcod") %>%
  #select(opcod, station, composition) %>%
  #unnest(composition) %>%
  #group_by(station, troph_group) %>%
  #summarise(richness_tot = length(unique(species)))
com_data <- compute_community_temporal_analysis(.op = op_analysis_bbb,
 type_network_metrics = "classes" 
)
# Compute sem dataset 
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]],
  hab_press = habitat_pressure,
  sync = com_data[["sync_std"]],
  nmds = NULL,
  basin = st_basin
  )

```{r, eval = FALSE}
stab_div_troph <- com_data$tps_bm_troph %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group),
    log_rich = log10(richness_med),
    log_rich_std = log10(rich_std_med),
    log_rich_tot = log10(richness_tot),
    log_rich_tot_std = log10(rich_tot_std),
    log_bm = log10(biomass_med),
    log_bm_std = log10(bm_std_med),
    log_stab = log10(biomass_stab),
    log_stab_std = log10(bm_std_stab)
    ) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med, richness_tot, log_rich, log_rich_std, log_rich_tot, log_rich_tot_std, log_stab, log_stab_std, log_bm, log_bm_std, biomass_med, bm_std_med, richness_tot, rich_tot_std) %>%
  dplyr::left_join(st_basin, by = "station")
```

@@

@@

pca_rotated <- compute_rotated_pca(
  .data = pca_pressure[, !names(pca_pressure) %in% "station"],
  naxis = 5 
)

prop_var_expl_pca <- pca_rotated$rotated$Vaccounted["Proportion Var", ] %>%
  map_dbl(., ~round(.x * 100))

tot_var_expl_pca <- round(sum(pca_rotated$rotated$Vaccounted["Proportion Var", ]) * 100) 
```

Habitat description were obtained from fish sampling data. We kept variables
describing the ecosystem size namely the stream width, depth and the ecosystem
position (altitude and slope). Physico-chemical variables were obtained from
naiades web portal [http://www.naiades.eaufrance.fr/](http://www.naiades.eaufrance.fr/).
We chose flow and temperature as physical variables to describe streams as flow
impose swimming constraints which filters fish species that can establish
[@rolls_scaling_2018, @tonkin_flow_2018, @gutierrez-canovas_contrasting_2013].
Water temperature is linked to physiological constraints namely thermal
tolerance and oxygen concentration tolerance [REF] since the level of water
saturation in oxygen is inversely linked to water temperature. We took 
Biochemical Oxygen Demand (BOD) for a proxy of enrichment in the ecosystem [REF].
BOD is also inversely linked to oxygen concentration in water.
The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD resp.). For each recording
station, we applied a moving window averaging (more details in supplementary
material) to get rid of seasonal effects and took the annual mean (i.e. one
value by year and by station. As the environmental recording stations were not
located at the same place than the fish stations, we interpolated environmental
data with spatial statistical network model [@isaak_applications_2014]
implemented in the `SSN` R package [@hoef_ssn:_2014]. The interpolation
procedure is detailed in supplementary material. For each fish station,
environmental variables were matched with the year of fish sampling (except for
water temperature for which the record began in 2006). Then, the average
temperature, flow, BOD, river width and depth were computed as well as
coefficient of variation ($CV = \sigma \mu^{-1}$). As habitat and environmental
variables are often colinear, a principal component analysis (PCA) was performed
on the scaled variables. The PCA explained `r tot_var_expl_pca`% of the
variance. An axis rotation was performed to maximize the representation of the
variables by the principal components (Fig. \@ref(fig:pca-rotated)). We kept
five axis, representing respectively the (1) river size (river width & depth),
(2) the altitude, slope and decreasing average temperature, (3) average
enrichment (BOD) and CV of flow, (4) river size CV (CV of river width and depth)
and (5) CV of enrichment (CV of BOD). Additionally, we logged the axis one, two
and three to normalize their distribution and we reversed the third axis to get
increasing average temperature with increasing values of the axis.

@@

@@

```{r mod-div-total, eval = FALSE}
stab_div_mod <- lmer(
  data = sem_data,
  control = ctrl_lmer,
  formula = log_stab_std ~ log_rich_tot_std + (1 | basin)
)
stopifnot(!isSingular(stab_div_mod, tol = 1e-05))
stab_div_boot <- bootMer(stab_div_mod, mySumm2, nsim = 100)
stab_div_ci <- confint(stab_div_boot)
#stab_div_colin <- car::vif(stab_div_mod) not use if one response
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

@@

```{r mod-div-troph, fig.show = "hide", eval = FALSE}
stab_troph_mod <- lmer(data = na.omit(stab_div_troph),
  formula = log_stab_std ~ log_rich_tot_std + troph_group + log_rich_tot_std:troph_group + (1 | basin)
  #control = ctrl
)

@@

## get coeff
#coef_troph <- map(list(stab = stab_troph_mod, bm = bm_troph_mod), fixed.effects)
```
```{r stab-div-troph, eval = FALSE}
stab_div_by_troph <- compute_stat_troph_rich(
  .data = stab_div_troph,
  resp = "log_stab_std"
)

bm_div_by_troph <- compute_stat_troph_rich(
  .data = stab_div_troph,
  resp = "log_bm_std"
)
troph_perm_test <- map(c(stab = "log_stab_std", bm = "log_bm_std"),
  ~test_troph(.data = stab_div_troph, var = .x)
  )
troph_perm_test$bm$obs
troph_perm_test$stab$obs
troph_perm_test$bm$p_val
troph_perm_test$stab$p_val
```

@@

We seeked to disantangle the unravel the links between
environment and community structure on both biomass stability and total biomass.
To investigate complex relationships, direct and indirect, between variables,
Structural Equation Modeling (SEM) is one promising way  
[@grace_structural_2008, @grace_causal_2014, @lefcheck_piecewisesem_2016,
@grace_integrative_2016]. SEM makes possible to set causal relationships between
variables. This step can be critical, as the estimated parameters can be
determined by both the choice of the variables and the links between them.
Therefore, the set of variables and links shoud be supported by theory and
litterature.

We built a metamodel (set of variables and links between them). Basically, we
tested the links between environmental conditions, community structure and
ecosystem properties namely biomass stability on total biomass. Theory in
ecology predicts that community structure influences stability. In turn,
environment influences community structure but also directly stability.

@@

To characterize environment, we retained five axis of PCA on habitat and
climatic variables which represent different types of forcing (constant and
variability). The two first axis (RC1 and RC2) were linked to the physical
structure and position of the river and the altitutinal gradient (+temperature).
The third axis (RC3) was linked to enrichment (Biological Demand in Oxygen) and
CV of annual flow. The fourth and fifth (RC4 and RC5) axis were associated to
environmental variability, respectively to CV of width and depth of the river
and CV of enrichment.

The community structure was characterized by total species richness, median
average weighted trophic level and median connectance. Total species richness
was the number of unique fish species that were observed at a given station over
the whole sampling. Both median connectance and average trophic level were just
computed as the median over the time series. For the details of the
computation of the trophic networks, see Network inference part.

@@

We linked environmental variables both to community structure and stability
components. Rooted in assembly theory, environment filters species that can
establish in a community and thus affect community structure. (+ add specific
hypothesis about physical and climate). (1) We hypothesized that enrichment
(characterized by DBO) and mean annual temperature will increase species turnover
[@friedpetersen_drivers_2020]. (2) We hypothesized environmental forcing can also
decrease biomass stability directly without going by community structure. First,
instability of habitat and environmental factors may in turn drive biomass
instability [@hansen_climate_2013]. Secondly, average environmental conditions
such as enrichment can increase directly CV by increasing species biomass
fluctuations [@tabi_warming_2019]. Increasing temperature was also hypothesized
to decrease stability through starvation resulting from high metabolic activity
[@tabi_warming_2019].

---
# Species richness to community attributes 
---
To reveal the mechanisms behind the relationship between species richness and
biomass stability, causal links were set from species richness to other
community attributes. In particular, species richness decreased connectance
following results from @thebault_stability_2010; @winemiller_must_1989. But
theoretically, some ecological networks can show the inverse relationship
[@winemiller_must_1989]. (6) We hypothesized also that species richness will
modulate average weighted trophic level because of food web assembly processes.
A community with more fish species can support higher trophic level. Finally, we
set up causal relationships from community attributes to stability components.

---
#  Community attributes to stability
---
We hypothesized that (7) species richness decreased synchrony between species
[@loreau_species_2008] and increased $\overline{CV_{sp}}$
[@tilman_biodiversity_2006]. (8a) We expected that average trophic level had a
positive or negative relationship with synchrony, bottom-up effects should
result in positive whereas top-down in negative relationships. We expected a (8b)
negative relationship between average trophic level and $\overline{CV_{sp}}$ if
predators stabilised prey fluctuations or if the increasing part of
biomass in the predators decreased $\overline{CV_{sp}}$ because of their higher
body mass (if mean increase faster than variance). (9a) Connectance was expected
to decrease or increase $\overline{CV_{sp}}$ and (9b) synchrony because of top
down effects or bottom up effects.

@@

library(DiagrammeR)
meta_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "meta_stab_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*30/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = c(from = 0, to = 7.5),
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2"),
  metamodel = TRUE
)
p_stab_sem_rich <- export_diagram(
  fit = stab_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "stab_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*30/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = c(from = 0, to = 7.5),
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2")
)

meta_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "meta_bm_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*45/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = c(from = 0, to = 7.5),
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2"),
  metamodel = TRUE
)
p_bm_sem_rich <- export_diagram(
  fit = bm_sem_rich,
  file_path = mypath("manuscript", "bef_stability", "bm_sem"),
  format = "png",
  width = 700*2,
  height = 900*2  - 900*2*45/100,
  arrows = "ortho",
  val_thd = 0.05,
  env_x_pos_range = c(from = 0, to = 7.5),
  env_y_pos = 0,
  env_y_sep = 1.5,
  order_env_node = c("RC4", "log_RC3", "RC5", "log_RC1", "log_RC2")
)
```

@@

@@

```{r sem-plot, fig.cap = sem_caption}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple
